---
title: "INFOH423 – Project: Hack my Ride"
author: "Team: B.Sara, B. Sid Ahmed, H. Ammar, E. Oussama"
output: pdf_document
params: 
       dt_hour: dt_hour
       dt_moments: dt_moments
       dt_type: dt_type
       stops_bxl: stops_bxl
---


# 0.Description initial dataset

### 0.1 Data loading schema

### 0.2 Data description

# Task 1: "Analyze vehicle speed"

### 1.1 Task pipeline schema


### 1.2 Pseudo-algorithm schema 

### 1.2 Let's have a look on the data and summary stat.
```{r, echo = FALSE, warning = FALSE}

# Modules to load 
library(data.table)
library(lubridate)
library(openxlsx)
library(DT)
library(ggplot2)
library(skimr)

# View Data 
datatable(dt_hour)

# Summary Stat of input 
skim(dt_hour)

```

### 1.3 Vizualize hourly average speed by typology
```{r, echo = FALSE, warning = FALSE, fig.width = 15}

metro <- ggplot(data = dt_hour[type_vehicle == 'm', ], aes(x = hour, y = avg_speed_hour, col = line_id, group = line_id)) + 
  geom_point() +
  geom_line() +
  theme_minimal() +
  ggtitle('What is the average hourly speed of the METRO STIB lines?')
metro

bus <- ggplot(data = dt_hour[type_vehicle == 'b', ], aes(x = hour, y = avg_speed_hour, col = line_id, group = line_id)) + 
  geom_point() +
  geom_line() +
  theme_minimal() +
  ggtitle('What is the average hourly speed of the BUS STIB lines?')
bus


tram <- ggplot(data = dt_hour[type_vehicle == 't', ], aes(x = hour, y = avg_speed_hour, col = line_id, group = line_id )) + 
  geom_point() +
  geom_line() +
  theme_minimal() +
  ggtitle('What is the average hourly speed of the TRAM STIB lines?')
tram

```


### 1.4 Visualize plots by moments and typology

```{r, echo = FALSE, warning=F, fig.width=15}
dt_moments[, line_id := as.factor(line_id)]
dt_moments[, moments := factor(moments, level = c('morning', 'afternoon', 'evening', 'night'))]

ggplot(data = dt_moments, 
       aes(x = moments, y = avg_speed_moments, col = type_vehicle, group = type_vehicle ))  +
  geom_point() +
  geom_line() + 
  theme_minimal() + 
  ggtitle('Average speed by moments: all lines')
```




# Task 2: "Analyze vehicle delays in Seconds."


### 2.1 Task pipeline schema


### 2.2 Pseudo-algorithm schema 


### 2.2. Visualization by vehicle type
```{r, echo = F, warning =F}
metro_delay <- ggplot(data = dt_hour[type_vehicle == 'm', ], aes(x = hour, y = avg_delay_hour, col = line_id, group = line_id)) + 
  geom_point() +
  geom_line() +
  theme_minimal() +
  ggtitle('What is the average hourly delay (sec.) of the METRO STIB lines?')
metro_delay

bus_delay <- ggplot(data = dt_hour[type_vehicle == 'b', ], aes(x = hour, y = avg_delay_hour, col = line_id, group = line_id)) + 
  geom_point() +
  geom_line() +
  theme_minimal() +
  ggtitle('What is the average delay (sec.) of the BUS STIB lines?')
bus_delay


tram_delay <- ggplot(data = dt_hour[type_vehicle == 't', ], aes(x = hour, y = avg_delay_hour, col = line_id, group = line_id )) + 
  geom_point() +
  geom_line() +
  theme_minimal() +
  ggtitle('What is the average hourly delay (sec.) of the TRAM STIB lines?')
tram_delay
```


### 2.3 Visualization by moments and vehicle 
```{r, echo = FALSE, warning = FALSE, fig.width=15}
ggplot(data = dt_moments, 
       aes(x = moments, y = avg_delay_moments, col = type_vehicle, group = type_vehicle))  +
  geom_point() +
  geom_line() + 
  theme_minimal() + 
  ggtitle('Average delay moments: all lines')
```


### 2.4 Testing coordinates map w.r.t to stops


### Data coordinates w.r.t to the stops 
```{r, echo =FALSE, warning =FALSE}
datatable(stops_bxl)
```

